---
title: "Data Science for Economists"
# subtitle: "<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>"
subtitle: "Lecture 13: Data Visualization - part 2, Maps"
author: "Drew Van Kuiken"
date: "University of North Carolina | [ECON 370](https://github.com/drewvankuiken/ECON370)" #"`r format(Sys.time(), '%d %B %Y')`"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#4B9CD3",
  header_font_google = google_font("Fira Sans","600"),
  text_font_google   = google_font("Fira Sans", "300", "300i"),
  base_font_size = "20px",
  text_font_size = "0.9rem",
  code_inline_font_size = "0.9em",
  code_font_size = "0.7rem",
  header_h1_font_size = "2.0rem",
  header_h2_font_size = "1.75rem",
  header_h3_font_size = "1.25rem",
  code_font_google   = google_font("Fira Code"),
  inverse_link_color = "#13294B",
  link_color = "#4B9CD3"
)
```


```{r setup, include=FALSE}
# xaringanExtra::use_scribble() ## Draw on slides. Requires dev version of xaringanExtra.

options(htmltools.dir.version = FALSE)
library(knitr)
opts_chunk$set(
  fig.align="center",  
  fig.height=4, #fig.width=6,
  # out.width="748px", #out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=T#, echo=F, warning=F, message=F
  )
library(tidyverse)
library(hrbrthemes)
library(fontawesome)
source("./lec13_maps_script.R")
```

# Maps

```{r, echo=FALSE}
ggplot(nyc_data) + 
  geom_sf(aes(geometry=geometry), colour = "white") +
  geom_point(aes(x=longitude, y=latitude, size=stump_diam), alpha=0.5) +
  scale_size_continuous(name = "Stump diameter") +
  labs(
    x = "Longitude", y = "Latitude",
    title = "Sample of New York City trees",
    caption = "Source: NYC Open Data"
  ) +
  theme_minimal()
```

`r fa('star-of-life')` Slides adapted from Grant McDermott's EC 607 at University of Oregon.

---
# Maps in R

Packages for today:
```{r}
## Load and install the packages that we'll be using today
if (!require("pacman")) install.packages("pacman")
pacman::p_load(sf, tidyverse, tigris, tidycensus, jsonlite, httr, maps, spData)
```

**New packages: sf, tigris, tidycensus**

---
# Features

We'll use the `sf` and `tigris` packages to get our map data today. Be forewarned: this is a quick and dirty exposition, enough to get you and up and running, but not a full treatment of maps. 

`sf`:
- Stands for *simple features*
- Treats features (e.g., zip codes, borders, trees, you name it) as data frames
- Great practice for data wrangling!

---
# Shapefiles

The sf package has some maps pre-loaded for us. Let's start by looking at the counties of North Carolina. 

```{r}
file_loc = system.file("shape/nc.shp", package="sf")
nc = st_read(file_loc, quiet = TRUE)
head(nc)
```

---
# Our first plot

```{r}
ggplot(nc) + geom_sf(aes(geometry=geometry, fill=AREA), col="white") + scale_fill_viridis_c(name = "Area") + ggtitle("Counties of North Carolina")
```
---
# Geodetic CRS

Otherwise known as the **projection** we're using. There are [lots](https://xkcd.com/977/) of projections we can use. 

To change projection, we can use the `st_transform` function:
```{r}
nc_moll <- nc |> 
  st_transform(crs = "+proj=moll") # using the Mollweide projection
head(nc_moll)
```

---
# Mollweide? 

Looks like this: 
```{r, echo=FALSE}
ggplot(nc_moll) + geom_sf(aes(geometry=geometry, fill=AREA), col="white") + scale_fill_viridis_c(name = "Area") + ggtitle("North Carolina in Mollweide")
```

Not super useful. Keep the projection thing in mind though, it will come up later. 

---
# Wrangling

We can use the tidyverse to work with sf objects:
```{r}
triangle <- nc |> 
  filter(NAME %in% c("Durham","Wake","Orange", "Chatham")) |> 
  mutate(AREA = AREA*1000) |> 
  select(NAME,AREA, everything())
head(triangle)
```

---
# Special Operations

We can melt geographies into larger geographic units using `st_union`:

```{r}
triangle |> st_union() |> ggplot() + geom_sf(fill=NA,col="black") + labs(title = "The Triangle") + theme_minimal()
```

---
# Other Operations

We're not going to focus on the full suite of operations available today. Here's a quick overview: 
- `st_area`: get area of shape
- `st_centroid`: get geographic center of area
- `st_boundary`: returns boundary of geometry
- `st_buffer`: creates buffer around each observation
- `st_distance`: distance between two objects (or more)

---
# Multiple Datasets

Let's pick up two new datasets:
```{r}
france = st_as_sf(map('france', plot=FALSE, fill=TRUE))
france = france[-c(95),] # observation causing some issues, not important
head(france)
data("seine", package = "spData")
head(seine)
```

---
# Merging

A couple thoughts: 
1. `st_as_sf`: turns non-shapefile data into shapefile data. `st_as_sf(x,coordinates=c("longitude","latitude"))` turns any lat/lon data into a sf object.
2. `st_intersect`: creates a dataset with the exact points of overlap between two objects. 
3. `st_join`: joins together two sf objects based on any overlapping geometries
  - Example of 2 and 3 in a sec
  
Try taking the intersection of the france and seine sf objects now.

---
# Didn't Work, Huh

Why didn't it work? 

--
<br>

```{r}
seine_crs = st_transform(seine, crs = st_crs(france))
france_intersected = st_intersection(france, seine_crs) 
```
The projections were different!

---
# So what'd we create?

```{r,echo=FALSE}
france_intersected |> ggplot() + geom_sf(alpha = 0.8, aes(fill = ID, col = ID)) + labs(title = "Seine, Marne and Yonne rivers", caption = "Colours depict French administrative regions") + theme(legend.title = element_blank())
```

---
# Trying out joins

```{r}
france_river <- st_join(france,seine_crs) |> # what kind of join is this?
  filter(!is.na(name)) |> 
  distinct(ID, .keep_all = T) # some rows merge twice because 2 branches of river
head(france_river)
```

---
# Put It In a Graph

```{r, echo=FALSE}
ggplot(france_river) + 
  geom_sf(alpha = 0.8, fill = "black", col = "gray50") + 
  geom_sf(data = seine, col = "#05E9FF", lwd = 1) + 
  labs(title = "Intersected regions only")
```

---
# tidycensus

The tidycensus package lets you avoid setting up even an API call to get census data, it's sort of amazing. 

We'll only scratch the surface of the tidycensus package. If you wanted to get ACS data (zipcode-level survey on all kinds of stuff), you'd need to grab an API key from the census. The instructions to do this are in the class script for today. 

One thing that's available in the tidycensus: TIGER/Line shapefiles (to access these, we loaded the `tigris` package)

Tiger shapefiles can give us shapefiles for things like: nations, regions, states, counties, census tracts, school districts, zip codes, and [way more](https://github.com/walkerke/tigris)

---
# Zip Codes in NY

Let's grab the zip codes for the urban area New York City:
```{r}
ny_zips <- zctas(state = "NY", class = "sf", year = 2010) # zips only available in 2010
urb <- urban_areas(year=2020) |> filter(grepl("New York",NAME10))
ny_urb_zips <- st_join(ny_zips,urb) |> filter(!is.na(NAME10))
```

Your task -- create 2 graphs:
1. Zip codes of NYC which contain trees downloaded from NYC Open Data
2. All zip codes in NY urban area, with points representing trees in NYC
