---
title: "Data Science for Economists"
# subtitle: "<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>"
subtitle: "Lecture 17: Presentation Tips"
author: "Drew Van Kuiken"
date: "University of North Carolina | [ECON 370](https://github.com/drewvankuiken/ECON370)" #"`r format(Sys.time(), '%d %B %Y')`"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#4B9CD3",
  header_font_google = google_font("Fira Sans","600"),
  text_font_google   = google_font("Fira Sans", "300", "300i"),
  base_font_size = "20px",
  text_font_size = "0.9rem",
  code_inline_font_size = "0.9em",
  code_font_size = "0.7rem",
  header_h1_font_size = "2.0rem",
  header_h2_font_size = "1.75rem",
  header_h3_font_size = "1.25rem",
  code_font_google   = google_font("Fira Code"),
  inverse_link_color = "#13294B",
  link_color = "#4B9CD3"
)
data(mtcars)
```


```{r setup, include=FALSE}
# xaringanExtra::use_scribble() ## Draw on slides. Requires dev version of xaringanExtra.

options(htmltools.dir.version = FALSE)
library(knitr)
opts_chunk$set(
  fig.align="center",  
  fig.height=4, #fig.width=6,
  # out.width="748px", #out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=T#, echo=F, warning=F, message=F
  )
library(tidyverse)
library(hrbrthemes)
library(fontawesome)
```
# Assorted Details

Final will be held in Gardner 307 - our normal classroom

To submit your final project: 
- Submit via **Dropbox**
- Folder link: [link](https://www.dropbox.com/scl/fo/r26aw4pii0a1uuatzp2fc/AALbvUF-ToHWAakyz2cU7I8?rlkey=w0k6ptkzje1l9me1x5sgygp0m&st=nja6x67a&dl=0). I'll also send this link out in a class announcement. 
- Create a subfolder titled firstname_lastname (no spaces!)
- Upload: PDF, HTML, and Rmd version of short slides and long slides
- If you've changed your project, please upload new data to sharepoint folder
- ***Due 3:30PM on Tuesday, December 10*** (i.e., 30 minutes before final period)

---
# In-class Presentations

- We have 3 hours and 30 presentations to do
- 5 minutes/presentation, 1 minute to do changeovers
  - Hence, no questions
  - In the real world, there are questions 
- Shoot for 1 minute/slide
  - I'll hold up a sheet of paper or something warning you when you have 1 minute left

---
class: inverse, center, middle

# Drew's Presentation Tips

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>

---
# Slides

- Slides are a visual reference, the things you say are what matter
--

- Make your slides in a way that feels natural
  - Go back over them and cut ~half the words out of your slides 
  - You still probably have too many words
--

- Keep your slides simple
  - 1 figure (graph or table)
  - Limited bullet points
  - Pretty is good but distracting is bad
--

- Slide titles
--

- Putting stuff on the slide iteratively can be helpful
--

- Or distracting
--

- Note: for ~20 slide presentation, this doesn't apply as much. Make sure it's legible for me and the grader, but more information is alright there

---
# Talking

- Public speaking is hard! 
- It is probably the most important skill you'll learn in college
--

- How would you describe your project if you had to explain it in a 5-minute elevator ride? 
  - Presentation is a story
--

- Easier to keep takeaways from a slide to 1-3 main ideas 
  - Some slides (Data) can be more: audiences wants to know details, doesn't need to keep track of them 
